### 2018-06-11 express2018/expansion.tex
Fixed and added basic definitions.

The equivalences here are 
 \emph{weak}, in that they
abstract from the number of internal steps performed
\begin{definition}[bisimilarity]
\label{d:wb}
A process relation ${\R}$ 
 is a \emph{bisimulation} if, whenever
 $P\RR Q$, %for all $\mu$ 
we have:
\begin{enumerate}
\item 
    $P \arr\mu P'$   implies that there is $Q'$ such that $Q \Arcap
\mu Q'$ and $P' 
\RR Q'$;

\item the converse of (1)
 on the actions from $Q$.
%\item if $Q  \Ar\mu   Q'$, then there is $P'$ s.t.\ $P
%\Arcap\mu  P'$ and $P' 
%\RR Q'$.
\end{enumerate}  
 $P$ and $Q$ are \emph{bisimilar},
written $P \wb
 Q$, if $P \RR Q$ for some  bisimulation $\R$.  
\end{definition}

\begin{definition}{rooted bisimilarity}

\end{definition}
\finish{the definition of rooted bisimilarity }

@@

\begin{definition}[expansion]
\label{d:expa}
A process relation ${\R}$ 
 is an {\em  expansion} if, whenever
we have $P\RR Q$, %for all $\mu$

\begin{enumerate}

\item   $P \arr\mu P'$ implies that there is $Q'$ with $Q \arcap \mu
  Q'$
 and $P' \RR Q'$;

\item 
    $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu
 P'$ and $P' 
\RR Q'$.

\end{enumerate}
 $P$  {\em expands } $Q$, written
$P  \contr Q$, 
if $P \RR Q$,  for some expansion $\R$. 
\qed\end{definition}

\finish{here the definition of contraction}

As bisimilarity, so expansion is preserved by all operators but sum.

>>>>

The equivalences here are 
 \emph{weak}, in that they
abstract from the number of internal steps performed
\begin{definition}[bisimilarity]
\label{d:wb}
A process relation ${\R}$ 
 is a \emph{bisimulation} if, whenever
 $P\RR Q$, %for all $\mu$ 
we have:
\begin{enumerate}
\item 
    $P \arr\mu P'$   implies that there is $Q'$ such that $Q \Arcap \mu Q'$ and $P' \RR Q'$;
\item the converse of (1)
 on the actions from $Q$.
%\item if $Q  \Ar\mu   Q'$, then there is $P'$ s.t.\ $P
%\Arcap\mu  P'$ and $P' 
%\RR Q'$.
\end{enumerate}  
 $P$ and $Q$ are \emph{bisimilar},
written $P \wb
 Q$, if $P \RR Q$ for some  bisimulation $\R$.  
\end{definition}

\begin{definition}{(rooted bisimilarity)}
Two processes $P$ and $Q$ are rooted bisimilar, denoted $P
\approx^c Q$, if for all $\mu$ (including $\tau$):
\begin{enumerate}
 \item  $P \arr\mu P'$   implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \approx Q'$;
\item the converse of (1) on the actions from $Q$.
\end{enumerate}
\end{definition}

@@

\begin{definition}[expansion]
\label{d:expa}
A process relation ${\R}$ 
 is an {\em  expansion} if, whenever
we have $P\RR Q$, %for all $\mu$ 
\begin{enumerate}
\item   $P \arr\mu P'$ implies that there is $Q'$ with $Q \arcap \mu
  Q'$
 and $P' \RR Q'$;
\item 
    $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu
 P'$ and $P' 
\RR Q'$.
\end{enumerate}
 $P$  {\em expands} $Q$, written
$P  \expa Q$, 
if $P \RR Q$,  for some expansion $\R$. 
\end{definition}

\begin{definition}[contraction]
\label{d:expa}
A process relation ${\R}$ is an {\em contraction} if, whenever
we have $P\RR Q$, %for all $\mu$ 
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ with $Q \arcap \mu Q'$
 and $P' \RR Q'$;
\item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arcap \mu
 P'$ and $P' \wb Q'$.
\end{enumerate}
 $P$  {\em contracts} $Q$, written
$P  \contr Q$, 
if $P \RR Q$,  for some contraction $\R$. 
\end{definition}

As bisimilarity, expansion and contraction are preserved by all operators but sum.

### 2018-06-11 express2018/unique.tex
Fixed and added basic definitions.

\finish{sopra: sum-free} 
\section{Rooted contraction and unique solution for rooted bisimilarity}
\label{ss:new}

\begin{definition}[rooted contraction]
\label{d:rcontra}
Two processes $P$ and $Q$ are in the rooted contraction, written 
 $P\rcontr Q$, 
if

\begin{enumerate}

\item   $P \arr\mu P'$ implies that there is $Q'$ with $Q \arr \mu
  Q'$
 and $P'\mcontrBIS Q'$;

\item 
    $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu
 P'$ and $P' 
\mcontrBIS Q'$.

\end{enumerate}
\qed\end{definition}

\finish{the only difference: the hat missing in the first clause}

\begin{theorem}
\label{t:}
Relation $\rcontr$ is a precongruence in CCS. 
\end{theorem}

@@

\begin{lemma}
\label{l:uptocon}
Suppose $\til P$ and $\til Q$ are solutions  for $\rcontr$ 
 of a system of weakly-guarded
contractions.
For any context $\qct$, 
if  $\ct{\til P}\Arr{\mu}  R$,
 then 
there is a  context $\qctp$
such that $R \mcontrBIS \ctp{\til P}$ and  $\ct{\til Q} \Arcap{\mu}
 \wb \ctp{\til Q}$.
\end{lemma}
\finish{sopra: si potrebbe togliere il hat, secondo me, e allora avremmo il risultato per
  rooted bisimilarity}

>>>>

\finish{sopra: sum-free}

\section{Rooted contraction and unique solution theorem for rooted bisimilarity}
\label{ss:new}

\begin{definition}[rooted contraction]
\label{d:rcontra}
Two processes $P$ and $Q$ are in the rooted contraction, written 
 $P\rcontr Q$, if
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ with $Q \arr \mu Q'$
 and $P'\mcontrBIS Q'$;
\item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu
 P'$ and $P' \wb Q'$.
\end{enumerate}
\end{definition}

\begin{theorem}
\label{t:rcontr-precongruence}
Relation $\rcontr$ is a precongruence in CCS. 
\end{theorem}

@@

\begin{lemma}
\label{l:uptocon}
Suppose $\til P$ and $\til Q$ are solutions  for $\rcontr$ 
 of a system of weakly-guarded
contractions.
For any context $\qct$, 
if  $\ct{\til P}\Arr{\mu}  R$,
 then 
there is a  context $\qctp$
such that $R \mcontrBIS \ctp{\til P}$ and  $\ct{\til Q} \Arcap{\mu}
 \wb \ctp{\til Q}$.
\end{lemma}
